---
title: "SIT741_Assignment 2"
author: "Shantanu Gupta"
date: "17 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Task 1:Source weather data

```{r}
library(magrittr) # need to run every time you start R and want to use %>%
library(dplyr)    # alternative, this also loads %>%
library(tidyverse)
library(tidyr) 
library(fitdistrplus)
library(scales)
library(rnoaa)#NOAA API
library(ISLR)
library(skimr)
library(mgcv)   ## GAM Package
library(gamair)
library(lubridate) #Dates and Time Package
library(broom)
```
```{r}
options(noaakey = "vlkCHTlbHxaMelPjFsikwjEjYqnXdHQT") #Token
Perth_data <- ncdc(
  datasetid = 'GHCND',
  stationid = 'GHCND:ASN00009021',
  startdate = '2013-07-01',
  enddate = '2014-06-30',
  limit = 1000
)$data
```


```{r}
a<-nrow(Perth_data)
sprintf("Rows in the dataset %s",a)
```

### Task 3:Model the ED demands

```{r}
ed_data_link <- 'govhack3.csv'
top_row <- read_csv(ed_data_link, col_names = FALSE, n_max = 1)
second_row <- read_csv(ed_data_link, n_max = 1)

column_names <- second_row %>% 
  unlist(., use.names=FALSE) %>% 
  make.unique(., sep = "__") # double underscore

column_names[2:8] <- str_c(column_names[2:8], '0', sep='__')

daily_attendance <- read_csv(ed_data_link, skip = 2, col_names = column_names)

daily_attendance

hospital_name<-top_row[colSums(!is.na(top_row)) > 0]
```

```{r}
daily_attendance$Date <- dmy(daily_attendance$Date)
daily_attendance <- transform(daily_attendance, ndate = as.numeric(Date),
                    nyear  = as.numeric(format(Date, '%Y')),
                    nmonth = as.numeric(format(Date, '%m')),
                    day    = as.numeric(format(Date, '%j')))
str(daily_attendance)
```

####Task 3.1: Models for a single facility
1.Pick a hospital
```{r}
sprintf("Royal Perth Hospital")
```

2.Linear Model
#####Linear Model
```{r}
lmfit_royal <- lm(Attendance__0~day , 
            data = daily_attendance) # Linear Model
lmfit_royal
```


```{r}
lmfit_royal %>% 
  summary  #Summary
```

```{r}
plot(lmfit_royal) #Plot
```

```{r}
plot(cooks.distance(lmfit_royal), pch = 16, col = "black")
```

```{r}
CD<-cooks.distance(lmfit_royal)
rev(sort(round(CD,5)))
```


```{r}
daily_attendance %>% 
  ggplot(aes(x = day, 
             y = Attendance__0)) +
  geom_point() +
  geom_smooth(method = 'lm') #Visualizing Linear Model
```

#####Poisson and General Linear Model
```{r}
poissonOut<-glm(Attendance__0~day , 
            data = daily_attendance,family="poisson")
poissonOut
```

```{r}
summary(poissonOut)
```

```{r}
tidy(poissonOut)
```

```{r}
coef(poissonOut)
```

```{r}
confint(poissonOut)
```

```{r}
ggplot(data=daily_attendance,aes(x = day, 
             y = Attendance__0))+
  geom_jitter(width=0.05,height=0.05)+
  geom_smooth(method='glm',method.args=list(family='poisson')) #Visualizing Poisson 
```

#####Negative Binomial and General Linear Model
```{r}
neg_binom_glm<-glm.nb(Attendance__0~day , 
            data = daily_attendance)
neg_binom_glm
```

```{r}
summary(neg_binom_glm)
```


3. General Additive Model
```{r}
ct1 <- gam(Attendance__0 ~ s(day),data=daily_attendance,method="REML")
ct1
```

```{r}
summary(ct1)
```

```{r}
coef(ct1) #Coefficients
```


```{r}
augment(ct1)
```

```{r}
tidy(ct1)
```

```{r}
glance(ct1)
```

```{r}
plot(ct1,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue") #Partial Effect Plot #Visualizing GAM
```

```{r}
plot(ct1,seWithMean = TRUE,shift=coef(ct1)[1]) #Transforming Standard Error
```

```{r}
gam.check(ct1) #Model Checking
```

```{r}
concurvity(ct1,full=TRUE) #Checking Concurvity
```


```{r}
results %>% 
  ggplot(aes(sample = .resid)) +
  geom_qq() + 
  geom_qq_line(col = 'steelblue')
```

4.Akaike Information Criteria(AIC) and Bayesian Information Criteria(BIC)
```{r}
AIC(ct1)
AIC(lmfit_royal)
BIC(ct1)
BIC(lmfit_royal)
```

####(Optional task) Task 3.2:Models for all hospitals
```{r}
ct2<- gam(Attendance__1 ~ s(day),data=daily_attendance,method="REML") #Fremantle Hospital
summary(ct2)
glance(ct2)
plot(ct2,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ct2)
```

```{r}
ct3<- gam(Attendance__2 ~ s(day),data=daily_attendance,method="REML") #Princess Margaret Hospital for Children
summary(ct3)
glance(ct3)
plot(ct3,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ct3)
```

```{r}
ct4<- gam(Attendance__3 ~ s(day),data=daily_attendance,method="REML") #King Edward Memorial Hospital for women
summary(ct4)
glance(ct4)
plot(ct4,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ct4)
```

```{r}
ct5<- gam(Attendance__4 ~ s(day),data=daily_attendance,method="REML") #Sir Charles Gairdner Hospital
summary(ct5)
glance(ct5)
plot(ct5,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ct5)
```

```{r}
ct6<- gam(Attendance__5 ~ s(day),data=daily_attendance,method="REML") #Armadale/kelmscott District Memorial Hospital
summary(ct6)
glance(ct6)
plot(ct6,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ct6)
```

```{r}
ct7<- gam(Attendance__6 ~ s(day),data=daily_attendance,method="REML") #Swan District Hospital
summary(ct7)
glance(ct7)
plot(ct7,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ct7)
```

```{r}
ct8<- gam(Attendance__7 ~ s(day),data=daily_attendance,method="REML") #Rockingham General Hospital
summary(ct8)
glance(ct8)
plot(ct8,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ct8)
```

```{r}
ct9<- gam(Attendance__8 ~ s(day),data=daily_attendance,method="REML") #Joondalup Health Campus
summary(ct9)
glance(ct9)
plot(ct9,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ct9)
```

### Task 4:Heatwaves and ED demands
#####Task 4.1:Measuring heatwave
```{r}
options(noaakey = "vlkCHTlbHxaMelPjFsikwjEjYqnXdHQT") #Token
Perth_data_avg <- ncdc(
  datasetid = 'GHCND',
  datatypeid = 'TAVG',
  stationid = 'GHCND:ASN00009021',
  startdate = '2013-07-01',
  enddate = '2014-06-30',
  limit = 1000
)$data
```


```{r}
library(RcppRoll)
E_H_F<-Perth_data_avg %>%
  mutate(t_3d_after=roll_meanr(lead(value),3)) %>%
  mutate(EHF=t_3d_after-30)
E_H_F[is.na(EHF)]=0
E_H_F
```

```{r}
Perth_data_avg %>% 
  mutate(date = ymd_hms(E_H_F$date)) %>% 
  ggplot(aes(x=date, 
             y=E_H_F$EHF)) + 
  geom_line(color='steelblue') + 
  ylab("Excess Heat Factor")
```

#####Task 4.2:Models with EHF
```{r}
ehf_gam<-gam(Attendance__0 ~ s(day)+s(E_H_F$EHF),data=daily_attendance,method="REML")
summary(ehf_gam)
glance(ehf_gam)
plot(ehf_gam,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(ehf_gam)
```

####(Optional task) Task 4.3:Extra weather features
```{r}
extra_gam<-gam(Attendance__0 ~s(day)+s(E_H_F$EHF)+s(E_H_F$value),data=daily_attendance,method="REML")
summary(extra_gam)
glance(extra_gam)
plot(extra_gam,rug=TRUE,residuals = TRUE,pch=1,cex=1,shade=TRUE,shade.col="steelblue")
gam.check(extra_gam)
```


